---
output:
  word_document: default
  html_document: default
---
```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(mice)
library(VIM)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(caret)
library(readr)
```
```{r}
parole <- read_csv("parole.csv")
Blood <- read_csv("Blood.csv")
```

```{r}
parole = parole %>%
  mutate(male = as_factor(male)) %>% 
  mutate(male = fct_recode(male, "Male" = "1", "Female" = "0")) %>%
  mutate(race = as_factor(race)) %>%
  mutate(race = fct_recode(race, "White" = "1", "Other" = "2")) %>%
  mutate(state = as_factor(state)) %>%
  mutate(state = fct_recode(state, "Other"="1","Kentucky"="2","Louisiana"="3","Virginia"="4")) %>%
  mutate(crime = as_factor(crime)) %>%
  mutate(crime = fct_recode(crime, "Other"="1","Larceny"="2","Drugs"="3","Driving"="4")) %>%
  mutate(multiple.offenses = as_factor(multiple.offenses)) %>%
  mutate(multiple.offenses = fct_recode(multiple.offenses, "Yes"="1","No"="0")) %>%
  mutate(violator = as_factor(violator)) %>%
  mutate(violator = fct_recode(violator, "Violated"="1","NonViolator"="0"))
```

```{r}
set.seed(12345)
parole_split = initial_split(parole,prob=0.70,strata = violator)
train = training(parole_split)
test = testing(parole_split)
```

```{r}
parole_recipe = recipe(violator ~ male+race+state+crime+multiple.offenses+age+time.served+max.sentence, train)

tree_model = decision_tree() %>%
  set_engine("rpart", model = TRUE) %>%
  set_mode("classification")
parole_wflow =
  workflow() %>%
  add_model(tree_model) %>%
  add_recipe(parole_recipe)

parole_fit = fit(parole_wflow, train)
```

```{r}
parole_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")
```

```{r}
tree = parole_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")
```

```{r}
rpart.plot(tree)
```
```{r}
fancyRpartPlot(tree,tweak=1.25)
```


I would classify them as a probable non-violator. I followed the tree to the right since their state was not kentucky, virginia, or other, then right again since multiple offenses was not equal to no, then left since their sentence was less than 13 years.

```{r}
parole_fit$fit$fit$fit$cptable
```

The optimal CP value is 0.0339

```{r}
set.seed(123)
folds = vfold_cv(train, v=5)
```

```{r}
parole_recipe2 = recipe(violator ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model2 = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

tree_grid = grid_regular(cost_complexity(),
                          levels = 25) 

parole_wflow2 = 
  workflow() %>% 
  add_model(tree_model2) %>% 
  add_recipe(parole_recipe2)

tree_res = 
  parole_wflow2 %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_res
```

```{r}
tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```

```{r}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```

The CP value that yields the optimal accuracy value is 0.1

```{r}
final_wf = 
  parole_wflow2 %>% 
  finalize_workflow(best_tree)
```

```{r}
final_fit = fit(final_wf, train)

tree = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

#fancyRpartPlot(tree, tweak = 1.5) 

```

My guess is that the root would be 100% accurate because it uses a CP that maximizes accuracy.

```{r}
Blood = Blood %>%
  mutate(DonatedMarch = as_factor(DonatedMarch)) %>%
  mutate(DonatedMarch = fct_recode(DonatedMarch, "Yes" = "1", "No" = "0"))
```

```{r}
set.seed(1234)
blood_split = initial_split(Blood,prob=0.70,strata = DonatedMarch)
train2 = training(blood_split)
test2 = testing(blood_split)
```

```{r}
blood_recipe = recipe(DonatedMarch ~ Mnths_Since_Last + TotalDonations + Total_Donated + Mnths_Since_First, train2)

tree_model3 = decision_tree() %>%
  set_engine("rpart", model = TRUE) %>%
  set_mode("classification")
blood_wflow =
  workflow() %>%
  add_model(tree_model3) %>%
  add_recipe(blood_recipe)

blood_fit = fit(blood_wflow, train2)
```

```{r}
blood_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")
```

```{r}
tree2 = blood_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")
```

```{r}
rpart.plot(tree2)
```
```{r}
fancyRpartPlot(tree2,tweak=1.5)
```

```{r}
set.seed(1234)
folds2 = vfold_cv(train2, v=5)
```

```{r}
blood_recipe2 = recipe(DonatedMarch ~., train2) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model4 = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

tree_grid2 = grid_regular(cost_complexity(),
                          levels = 25) 

blood_wflow2 = 
  workflow() %>% 
  add_model(tree_model4) %>% 
  add_recipe(blood_recipe2)

tree_res2 = 
  blood_wflow2 %>% 
  tune_grid(
    resamples = folds2,
    grid = tree_grid2
    )

tree_res2
```

```{r}
tree_res2 %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```

```{r}
best_tree2 = tree_res2 %>%
  select_best("accuracy")

best_tree2
```


```{r}
final_wf2 = 
  blood_wflow2 %>% 
  finalize_workflow(best_tree2)
```

```{r}
final_fit2 = fit(final_wf2, train2)

tree2 = final_fit2 %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

#fancyRpartPlot(tree2, tweak = 1.5) 

```

```{r}
treepred = predict(final_fit2, train2, type = "class")
head(treepred)
```


```{r}
confusionMatrix(treepred$.pred_class,train2$DonatedMarch,positive="Yes") 
```

Predictions on testing set  
```{r}
treepred_test = predict(final_fit2, test2, type = "class")
head(treepred_test)
```

```{r}
confusionMatrix(treepred_test$.pred_class,test2$DonatedMarch,positive="Yes") 
```

On both the Testing and Training sets the model was about 76% accurate.