---
output:
  word_document: default
  html_document: default
---
# Random Forests
## Andrew Goff

```{r, include = FALSE}
library(tidyverse)
library(tidymodels)
library(mice) 
library(VIM) 
library(ranger) 
library(randomForest) 
library(caret)
library(skimr)
library(GGally)
library(gridExtra)
library(vip)
```

```{r}
drug = read_csv("drug_data-1.csv")
```

```{r}
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity","Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive","SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis","Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh","LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")

#str(drug)
```

```{r}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"
```

```{r}
drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44","45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels =c("Under16", "At16", "At17", "At18", "SomeCollege","ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
  mutate(Country = factor(Country,labels = c("USA", "NewZealand", "Other", "Australia","Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity,labels = c("Black", "Asian", "White", "White/Black", "Other","White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
  select(-ID)
```

```{r}
#str(drug_clean)
drug_clean = drug_clean %>% 
  select(!(Alcohol:Mushrooms)) %>% 
  select(!(Semer:VSA))

names(drug_clean)
```

```{R}
skim(drug_clean)

```

I don't see any missing data according to this chart.

```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) 
train = training(drug_split)
test = testing(drug_split)
```

Visualization  
```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```

```{r}
p1 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Nscore, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Escore, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Oscore, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```

```{r}
p1 = ggplot(train, aes(x = Ascore, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Cscore, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Impulsive, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = SS, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```

It appears younger males are more likely to use nicotine, impulsive and SS are also predictors I think.

Set up our folds for cross-validation  
```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```

```{r}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% #
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid
)
```

```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

```{r}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```

```{r}
final_rf_fit = fit(final_rf, train)
```

```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

It's apparent than SS is most important to the model, followed by Country (UK) and openness.  After that Female, age 45-54, and impulsiveness look to be similarly important.

```{r}
trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)
```


```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```


```{r}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

The model appears about 84% accurate on the training set and 71% accurate on the testing set.  

This model could be used in the real world to help target at risk populations for nicotine use/abuse.  The accuracy for this model is relatively good and in this case, targeting the wrong groups wouldn't really do much harm as anyone can and should be warned about nicotine abuse.